Manuscript draft

### 2019-06-04 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
updated manuscript draft

@@

Globally, forests play an important role in the carbon cycle. Old growth forests alone are estimated to sequester up to 1.4 GtCyr ^-1^ [@Malhi1999], making them a significant carbon sink, and an important component of global carbon dioxide budgets [@Luyssaert2008]. Understanding the patterns and mechanisms of forest carbon cycling on a global scale is therefore of high importance in understanding the potential role of forests in mitigating climate change.

Little is known about the main environmental drivers of these patterns.

The productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood. Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.

In order to better understand and predict responses and feedbacks to climate change, we need a better fundamental understand of how forest carbon cycling varies with respect to climate globally.

@@

*Understanding components of primary productivity.* There are several different measures of primary productivity. Gross primary productivity (GPP) represents the gross uptake of carbon dioxide by plants in an ecosystem, to be used in photosynthesis and energy production. Only a fraction of this carbon is assimilated into plant tissues; the rest is used in autotrophic respiration. The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem; thus GPP can be described as the sum of NPP and ecosystem respiration.

Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis. Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@Clark2017].

@@

@@

Data that control for stand age and standardize methodologies are required to resolve this question. Here, we use a comprehensive global database to explore global patterns in productivity.

>>>>

add hypotheses, can number/bullet point them
redo first couple paragraphs
look again at breakdown of primary productivity components --> add discussion on understanding belowground productivity

@@

Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@Luyssaert2008]. They show higher levels of productivity than non-forest  terrestrial ecosystems [@DelGrosso2008], and as a result achieve significant carbon sequestration and storage. Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@Malhi1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@Pan2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change. Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.

There are two major questions to understand: firstly, how forest productivity varies globally, and specifically which areas show the greatest peaks in productivity; and secondly, which climate variables drive this variation. On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@Beer2010; @Jung2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood. This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above. Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@DelGrosso2008], but doesn't necessarily respond linearly to these drivers. Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.

@@

*Understanding components of primary productivity.* There are several different measures of primary productivity. Gross primary productivity (GPP) represents the gross uptake of carbon dioxide by plants in an ecosystem, to be used in photosynthesis and energy production. Only a fraction of this carbon is assimilated into plant tissues; the rest is used in autotrophic respiration. The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem; thus GPP can be described as the sum of NPP and ecosystem respiration.   
Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis. Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@Clark2017].

@@

[[Belowground productivity is a potentially often underestimated component of productivity, but has been shown to be significant, particularly in higher latitude forests [@Pan2011].]]

@@

Data that control for stand age and standardize methodologies are required to resolve this question. Here, we use a comprehensive global database to explore global patterns in productivity. We explore three questions: 
1. Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?
2. Do the components of primary productivity show variation in their responses to these climatic drivers? Is carbon allocation climate invariant?
3. Does climate explain the same proportion of variation in different components of primary productivity? Is the relative importance of climatic drivers the same across components of primary productivity?
